Fusion Moves for Correlation Clustering

### 2014-11-11 paper/introduction-new.tex

\documentclass[10pt,twocolumn,letterpaper]{article}

\usepackage{eso-pic}
\usepackage{cvpr}
\usepackage{times}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{xspace}
\usepackage{booktabs}
\usepackage{placeins}

% Include other packages here, before hyperref.
\usepackage{my_macros}
\usepackage{paralist}
\usepackage{amsthm}
% \usepackage{dirtytalk} fah: my miktex didn't like this; and we don't need it here. 
%\usepackage{framed}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{algorithm}
%\usepackage{algorithmic}
\usepackage{algpseudocode}
\usepackage{setspace}
%for tikz
\usepackage{tikz,pgfplots}
\usepackage{pgfplotstable}
\usepackage{filecontents}
%\usepackage{scrextend}
\usepackage[pagebackref=true,breaklinks=true,letterpaper=true,colorlinks,bookmarks=false]{hyperref}

\newcommand{\footlabel}[2]{%
    \addtocounter{footnote}{1}%
    \footnotetext[\thefootnote]{%
        \addtocounter{footnote}{-1}%
        \refstepcounter{footnote}\label{#1}%
        #2%
    }%
    $^{\ref{#1}}$%
}

\newcommand{\footref}[1]{%
    $^{\ref{#1}}$%
}
\usetikzlibrary{arrows,positioning,automata,shadows,fit,shapes}
\usetikzlibrary{arrows,petri,topaths}
\usetikzlibrary{positioning,fit,calc}
\usetikzlibrary{shapes.arrows,chains,decorations.pathreplacing,fadings}
\usetikzlibrary{calc, matrix, backgrounds}

%%%%%%global tikz styles

\tikzstyle{eBlack}=[very thick,minimum size = 0.5cm,inner sep = 0pt,draw=black!70!black, font=\small,align=left, line width=1mm]
\tikzstyle{eNotCut}=[very thick,minimum size = 0.5cm,inner sep = 0pt,draw=green!70!black, font=\small,align=left, line width=1mm]
\tikzstyle{eBlue}=[very thick,minimum size = 0.5cm,inner sep = 0pt,draw=blue!70!black, font=\small,align=left, line width=1mm]
\tikzstyle{eCut}=[very thick,dotted,minimum size = 0.5cm,inner sep = 0pt,draw=black, font=\small,align=left, draw=red!70!black, line width=1mm]

\pgfplotsset{every axis/.append style={
  every axis y label/.style = {at={(ticklabel cs:0.5)}, rotate=90, anchor=south},
  axis x line = {bottom},
  axis y line = {left},
  tick align = outside,
  ymajorgrids = true,
%  legend style = {draw=none, at={(1.05, 0.5)}, anchor=west, font=\small},
  legend style = {font=\tiny},
  legend columns = 1,
  every axis plot/.append style = {line width=1pt},
  label style = {font=\small},
  tick label style={font=\small},
  scaled ticks = false,
}}

% Two Colored Circle Split 
\makeatletter
\tikzset{circle split part fill/.style  args={#1,#2}{%
 alias=tmp@name, 
  postaction={%
    insert path={
     \pgfextra{% 
     \pgfpointdiff{\pgfpointanchor{\pgf@node@name}{center}}%
                  {\pgfpointanchor{\pgf@node@name}{east}}%            
     \pgfmathsetmacro\insiderad{\pgf@x}
      \fill[#1] (\pgf@node@name.base) ([xshift=-\pgflinewidth]\pgf@node@name.east) arc
                          (0:180:\insiderad-\pgflinewidth)--cycle;
      \fill[#2] (\pgf@node@name.base) ([xshift=\pgflinewidth]\pgf@node@name.west)  arc
                           (180:360:\insiderad-\pgflinewidth)--cycle;            
         }}}}}  
 \makeatother

%\usepackage{tkz-berge}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\argmax}{arg\,max}
\definecolor{shadecolor}{rgb}{0.01,0.199,0.1}
\usepackage{xargs} 
\newtheorem{theorem}{Theorem}
\newtheorem{remark}{Remark}

% If you comment hyperref and then uncomment it, you should delete
% egpaper.aux before re-running latex.  (Or just hit 'q' on the first latex
% run, let it finish, and you should be clear).
\usepackage{bm}% ändert \boldsymbol
\def\arraystretch{0.8}
\renewcommand{\tabcolsep}{2pt}
\newcommand{\thickline}{2pt}
\newcommand{\scatterplotpath}{./scatterplots/}

\usepackage[colorinlistoftodos,prependcaption,textsize=tiny]{todonotes}
\newcommandx{\unsure}[2][1=]{\todo[linecolor=red,backgroundcolor=red!25,bordercolor=red,#1]{#2}}
\newcommandx{\change}[2][1=]{\todo[linecolor=blue,backgroundcolor=blue!25,bordercolor=blue,#1]{#2}}
\newcommandx{\info}[2][1=]{\todo[linecolor=OliveGreen,backgroundcolor=OliveGreen!25,bordercolor=OliveGreen,#1]{#2}}
\newcommandx{\improvement}[2][1=]{\todo[linecolor=Plum,backgroundcolor=Plum!25,bordercolor=Plum,#1]{#2}}
\newcommandx{\thiswillnotshow}[2][1=]{\todo[disable,#1]{#2}}
\newcommand{\OR}{\textrm{ or }}

% \cvprfinalcopy % *** Uncomment this line for the final submission

\def\cvprPaperID{****} % *** Enter the CVPR Paper ID here
\def\httilde{\mbox{\tt\raisebox{-.5ex}{\symbol{126}}}}

\newcommand{\realplot}[1]{
\begin{tikzpicture}
    \begin{axis}
        \addlegendimage{empty legend}\addlegendentry{Matrix #1}
        \addplot {0};
        \addplot {1};
        \addplot {2};
        \addplot {3};
    \end{axis}
\end{tikzpicture}}

% \newcommand{\anytimeplot}[1]{
%   \addplot[color=brown,mark=x] table[x=time,y=HC]{data3/knott-3d-300.data};              \addlegendentry{HC}
%   \addplot[color=black,mark=x] table[x=time,y=HC-CGC]{data3/knott-3d-300.data};          \addlegendentry{HC-CGC}
%   \addplot[color=red,mark=x] table[x=time,y=CGC]{data3/knott-3d-300.data};               \addlegendentry{CGC}
%   \addplot[color=gray,mark=o] table[x=time, y=ogm-KL]{data3/knott-3d-300.data};          \addlegendentry{KL}
%   \addplot[color=purple,mark=x] table[x=time, y=MCR-CCFDB]{data3/knott-3d-300.data};     \addlegendentry{MC-R}
%   \addplot[color=blue,mark=x] table[x=time, y=MCI-CCIFD]{data3/knott-3d-300.data};       \addlegendentry{MC-I}
%   \addplot[color=green,mark=x] table[x=time, y=DYNCC-HC-MC]{data3/knott-3d-300.data};    \addlegendentry{Fusion-HC-MC}
%   \addplot[color=cyan,mark=x] table[x=time, y=DYNCC-HC-CGC]{data3/knott-3d-300.data};    \addlegendentry{Fusion-HC-CGC}
%   \addplot[color=orange,mark=x] table[x=time, y=DYNCC-WS-MC]{data3/knott-3d-300.data};  \addlegendentry{Fusion-WS-MC}
%   \addplot[color=orange,mark=x] table[x=time, y=DYNCC-WS-CGC]{data3/knott-3d-300.data};  \addlegendentry{Fusion-WS-CGC}
% }
\newcommand{\anytimeplot}[1]{
  \addplot[color=brown,mark=x] table[x=time,y=HC]{data3/#1.data};              \addlegendentry{HC}
  \addplot[color=black,mark=square] table[x=time,y=HC-CGC]{data3/#1.data};          \addlegendentry{HC-CGC}
  \addplot[color=red,mark=x] table[x=time,y=CGC]{data3/#1.data};               \addlegendentry{CGC}
  \addplot[color=gray,mark=o] table[x=time, y=ogm-KL]{data3/#1.data};          \addlegendentry{KL} 
  \addplot[color=yellow!50!black,mark=square] table[x=time, y=ogm-mcfusion-HC-CF*]{data3/#1.data};  \addlegendentry{Fusion}
  \addplot[color=purple,mark=o] table[x=time, y=MCR-CCFDB]{data3/#1.data};     \addlegendentry{MC-R}
  \addplot[color=blue,mark=o] table[x=time, y=MCI-CCIFD]{data3/#1.data};       \addlegendentry{MC-I}
  \addplot[color=green,mark=o] table[x=time, y=DYNCC-HC-MC]{data3/#1.data};    \addlegendentry{CCFusion-HC-MC}
  \addplot[color=cyan,mark=x] table[x=time, y=DYNCC-HC-CGC]{data3/#1.data};    \addlegendentry{CCFusion-HC-CGC}
  \addplot[color=orange,mark=o] table[x=time, y=DYNCC-WS-MC]{data3/#1.data};  \addlegendentry{CCFusion-WS-MC}
  \addplot[color=pink,mark=x] table[x=time, y=DYNCC-WS-CGC]{data3/#1.data};  \addlegendentry{CCFusion-WS-CGC}
}
%  \addplot[color=yellow!50!black,mark=o] table[x=time, y=ogm-mcfusion-HC-CF*]{data3/#1.data};  \addlegendentry{Fusion}
  %\addplot[color=green!50,mark=o] table[x=time, y=PIVOT*]{data3/#1.data};          \addlegendentry{PIVOT-BOEM}

% Pages are numbered in submission mode, and unnumbered in camera-ready
\ifcvprfinal\pagestyle{empty}\fi
\begin{document}
%%%%%%%%% TITLE
\include{my_macros.tex}
%\title{Correlation Clustering with Dynamic Super-Nodes (DySNCC)}
\title{Fusion Moves for Correlation Clustering}

\author{First Author\\
Institution1\\
Institution1 address\\
{\tt\small firstauthor@i1.org}
% For a paper whose authors are all at the same institution,
% omit the following lines up until the closing ``}''.
% Additional authors and addresses can be added with ``\and'',
% just like the second author.
% To save space, use either the email address or home page, not both
\and
Second Author\\
Institution2\\
First line of institution2 address\\
{\tt\small secondauthor@i2.org}
}

\maketitle
%\thispagestyle{empty}

%%%%%%%%% ABSTRACT
\begin{abstract}
  \color{red} \textbf{TODO:}
Correlation clustering, or multicut partitioning, is the NP-hard problem of partitioning an undirected graph or image with positive / attractive and negative / repulsive edge weights such that the sum of cut edge weights is minimized. We here present a move-making heuristic that uses cheaply generated proposal clusterings to iteratively and monotonously improve upon the best previously found clustering. A valid partitioning is maintained at all times, and solutions close to the global optimum are obtained faster than with any published technique.

%We address the problem of partitioning a  graph
%   into a previously unknown number of clusters.
%   Among all partitions, the one with the minimal 
%   sum of of cut edge weights is chosen. 
%   This problem is known as correlation clustering 
%   or multicut.
%   We propose a general framework to find
%   high quality approximate solutions for 
%   this NP-hard problem based on a move making algorithm:
%   We use candidate solutions from a proposal generator
%   to iteratively improve the best observed solution similar
%   to fusion moves.
%   The proposed solver outperforms any other solver
%   w.r.t. to any time performance.
%   The solutions found by this solver are close
%   to global optimal solutions w.r.t. energy
%   and problem specific measurements as VI for
%   image segmentation problems.

\end{abstract}
\section{Introduction}

Correlation clustering~\cite{Bansal-2002} also known as the multicut problem~\cite{chopra_1993_mp} 
is a basic primitive in computer vision~\cite{andres_2011_iccv,kroeger_2012_eccv,yarkony_2012_eccv,alush_2013_simbad} and data mining~\cite{Chierichetti-2014,Arasu-2009,Sadikov-2010,Chen-2012}. Its value is, firstly, that it accommodates both positive / attractive \emph{and} negative / repulsive edge weights. Importantly, this allows doing justice to evidence in the data that two nodes or pixels do \emph{not} wish to end up in the same cluster or segment. This is in contrast to the submodular potentials so popularized by the graph cut algorithm, which only allow two nodes to be attracted to each other, or at most be agnostic about membership in the same cluster. Secondly, the algorithm does not require a specification of the number of clusters. This is in contrast to methods such as normalized cut \cite{} that can only accommodate attractive interactions and hence need the number of clusters to be specified.

\paragraph{Contribution.} The evident usefulness of correlation clustering and its clean and compact formulation in terms of an optimization problem, together with its unfortunate NP-hardness, are an invitation to develop fast approximate solvers. The basic idea of the move making algorithm proposed here is to maintain, at all times, a best current partitioning; and to iteratively improve it (as we show, monotonously) by considering diverse and cheaply generated proposal partitionings. Any two nodes that are in one cluster in both the current and the proposal partitioning are contracted. Edges to contracted nodes are equally combined and reweighted, and the correlation clustering problem is then solved on this reduced graph. We offer a polyhedral characterization of this strategy, and evaluate it in conjunction with two versatile proposal generators.   We conduct experiments on a broad range of data sets ranging from 2D and 3D segmentation problems to clustering in signed social networks. The results suggest that this simple move making algorithm is the fastest technique known today, reaching close to globally optimal solutions in a tenth or a hundredth of the time required by the most efficient exact solvers.

\paragraph{Related Work.}
A natural approach is to solve the integer linear programming problem directly. To this end, efficient separation procedures have been found \cite{} that allow to iteratively augment the set of constraints until a valid partitioning is found. 
Alternatively, it is possible to relax the integrality constraints of the integer linear programming formulation. Such an outer relaxation can be iteratively tightened. However, intermediate solutions are fractional and so rounding is required to obtain a valid partitioning. The latter approach works best on planar \cite{yarkony_2012_eccv} or almost planar \cite{yarkony-andres-2013} graphs.

@@

---------------END OF SUGGESTION ----------------------------------

Correlation clustering~\cite{Bansal-2002} also known as the multicut problem~\cite{chopra_1993_mp} 
is a basic primitive in computer vision~\cite{andres_2011_iccv,kroeger_2012_eccv,yarkony_2012_eccv,alush_2013_simbad} and data mining~\cite{Chierichetti-2014,Arasu-2009,Sadikov-2010} as a first step
towards understanding an image or data in general. 
The correlation clustering problem is the following:
Given an undirected edge weighted graph, 
where positive weights (\emph{attractive}) encourage the adjacent nodes to be in the same cluster
and negative weights (\emph{repulsive}) encourage the nodes to stay in different cluster, 
find the clustering of nodes such that the sum over weights of edges between clusters is minimized.
Note that this problem formulation does not included a predefined number of clusters.

Despite the clear mathematical formulation and nice properties of correlation clustering
to perform a one-shot agglomeration clustering,
the main draw back of correlation clustering is scalability~\cite{nunez_iglesias_2013}
which arise from the NP-hardness of the problem~\cite{Bansal-2002}.

Recently, Beier~\etal~\cite{beier_2014_cvpr} presented a promising 
approximative method called Cut, Clue \& Cut.
While this scales well for planar graphs, it does not for non planar graphs and has a bad any-time performance.
A further inherent drawback of this method is that the subproblems are not limited in their size.
%
Consequently, partition problems on large scale data, \eg
huge volume images in computational neuroscience~\cite{kroeger_2012_eccv}
or social networks~\cite{Leskovec-2010}, 
are not tractable, because the runtime explodes or the 
approximations are far away from being useful.

In the present work we will suggest some alternative approaches to deal with large scale correlation clustering problems.
In the end we ill present a method that can be combined with any other correlation clustering method and 
allows to reduce the overall problem to a sequence of correlation clustering problems which are by magnitudes smaller.

\textbf{Contribution:}
In the present work we present some novel approaches that are designed for large scale correlation clustering problems.
First, we define a novel energy based agglomerative clustering algorithm that monotonically increase the energy.
With this at hand we show how to improve the anytime performance of Cut, Clue \& Cut.
Second, we introduce cluster-fusion moves, which extend the original fusion moves~\cite{Lempitsky-2010} 
used in supervised segmentation to the unsupervised case and give a polyhedral interpretation of this algorithm.
We propose two versatile proposal generators, and evaluate the proposed methods on existing and new benchmark problems.
That experiments show that we can improve the computation time by one to two magnitudes without worsening the segmentation 
quality significantly.

% Given an edge weighted graph, positive weights (\emph{attractive})
% encourage the adjacent nodes to be in the same 
% connected component, while negative weights (\emph{repulsive}) encourage
% the nodes to stay in different connect components.
% Therefore the sign of the weights encodes if two
% nodes should be merged or not and the magnitude of the weights encodes
% the certainty of this desire.
% The objective of correlation clustering / multicuts
% is finding the cut with a minimal sum of cut edge weights.
% The number of connected components / clusters is discovered
% from the weights ( rewrite?!?).
% Correlation clustering / the multicut is NP-hard \cite{???}.
% %
% Despite the NP-hardness, correlation clustering has been 
% successfully used for
% \begin{inparaenum}[(i)]
%     \item partitioning a superpixel region adjacency graph~\cite{andres_2011_iccv,kroeger_2012_eccv}
%     \item with optional long range repulsive edges~\cite{andres_2013_emmcvpr}.
%     \item Alush and Globerger showed how to average multiple segmentations with the multicut objective~\cite{alush_2012_pami}.
%     \item Multicuts can also be used for interactive segmentation~\cite{bagon_2011_arxiv},
%     \item for co-segmentation~\cite{glassner_2011_cvpr}
%     \item and to cluster sparse and graphs~\cite{???}.
% \end{inparaenum}

% Despite the nice properties of correlation clustering
% to perform a \say{one-shot agglomeration of supervoxels} ~\cite{nunez_iglesias_2013},
% the main draw back of correlation clustering is scalability~\cite{nunez_iglesias_2013}.
% Existing solvers for correlation clustering do not scale for non planar graphs.
% Even state of the art approximate move making algorithm fail 
% to give good approximations non planar graphs.
% When increasing the problem sizes, either the runtime explodes or the 
% approximations are far away from being useful.

% \textbf{Contribution:}
% \begin{inparaenum}[(i)]
% \item Within this work we propose a fast and scalable move making algorithm for correlation clustering,
% \item which generalized  fusion moves \cite{???} to correlation clustering problems.
% \item We reduce the inference to a series of moves where
% each moves optimizes over a subspace spanned by the current best solution
% and a proposal solution.
% \item We propose two versatile proposal generators,
% \item and show how to optimize these moves.
% \item We give a polyhedral interpretation of this algorithm,
% \item and evaluate the proposed method 
% on existing benchmark problems.
% \item We show state of the art any time performance on those instances.
% \end{inparaenum}[(i)]

\textbf{Outline:} {\bf If we are short of space, the following can be ommited:}
In sec.~\ref{sec:problem_formulation} we will give a 
detailed problem definition where we introduce 
the correlation clustering objective.
In sec.~\ref{sec:related_work} we will 
discuss existing methods for correlation 
clustering and briefly explain the concept of fusion moves.
In sec: ~\ref{sec:cc_fm} we describe our proposed
method and show the properties of the algorithm.
In sec.~\ref{sec:exp} we show an evaluation
of the method on existing parameter  and discuss the effects of parameters.
Future work will be discussed in sec. \ref{sec:future} and
we will conclude in ~\ref{sec:conclusion}.
%-------------------------------------------------------------------------

\section{Notation and Problem Formulation}\label{sec:problem_formulation}
Let $G=(V,E, w)$ be a weighted graph of nodes $V$ and edges $E$.
%
The function $w : E \rightarrow \mathbb{R}$ assigns a weight to each edge.
A positive weight expresses the desire that two adjacent nodes should
be merged, whereas a negative weight indicates
that these nodes should be separated into two different regions.
%
A \emph{subgraph} $G_A = \{A, E_A, w\}$ consists
of nodes $A \subseteq V$ and edges $E_A := E\cap (A\times A)$.
%
A segmentation of the graph $G$ can by either given by an 
node labeling $l \in \mathbb{N}^{|V|}$
or an edge labeling $y \in\{0,1\}^{|E|}$.  
An edge labeling is only consistent if it contains no dangling edges~\cite{kappes_2013_arxiv}.
We denote the set of all consistent edge labelings by $P(G)\subset\{0,1\}^{|E|}$.
The convex hull of this set is known as the \emph{multicut polytope} $MC(G) = \textrm{conv}(P(G))$.

%$\rho_y : P(G) \to P(G_y)$
%$\rho_y^{-1} : P(G_y) \to P(G)$

% \begin{itemize}
%   \item Multicut in computer vision, applications, sparse segmentation -> IMPORTANT PROBLEM
%   \item Problem: Existing Methods does not scale, even CGC. - Large scale 3d problems
%   \item Contribution: Fast scalable method using novel fusion moves for correlation clustering
% \end{itemize}

% Def: Graph, weighted graph, cut, multicut

Given a weighted graph $G=(V,E,w)$ we consider the problem of segmenting $G$ such that the costs
of the edges between distinct segments is minimized. This can be formulated in the node domain
by assigning each node $v$ a label $l_v \in \mathbb{N}$
\begin{align}
  l^* &= \argmin_{L \in \mathbb{N}^{|V|}} \sum_{ (i,j) \in E } w_{ij} \cdot [l_{i} \neq l_{j}], \label{eq:nodeproblem}
\end{align}  % y_{ij}^* &=& [l_{u} \neq l_{v}] 
or in the edge domain, by label each edge $e$ as cut $y_e=1$ or uncut $y_e=0$ 
\begin{align}
  y^* &= \argmin_{y \in P(G)} \sum_{ (i,j) \in E } w_{ij} \cdot y_{ij} \label{eq:edgeproblem}.%\\ 
\end{align}
As shown in~\cite{kappes_2013_arxiv} both problems are equivalent, but formulation \ref{eq:nodeproblem}
suffers from ambiguities in the formulation~\cite{kappes_2011_emmcvpr}.

% The exist an surjective mapping from a node-label $l$ to edge-labeling $y$ and
% a bijective mapping from a partitioning of $V$ to vertices of the multicut polytope $MC$.
% Consequently, 
% (i) problems \ref{eq:nodeproblem} and \ref{eq:edgeproblem} are equivalent
% and (ii) the node-labeling is not unique for a given partitioning and introduce some
% ambiguities.

% \subsubsection{Multicut Objective}

% The multicut / correlation clustering objective 
% can be formulated in different ways.

% \paragraph{Edge Indicator Variables:}
% \begin{center}
%     \begin{eqnarray}
%         y^* &=& \argmin_{y} \sum_{ e_{ij} \in E } w_{ij} \cdot y_{ij} \\
%         s.t.:& & y \in \textit{Multicut Polytope} \nonumber
%     \end{eqnarray}
% \end{center}

% \paragraph{Fully Connected Graph:}
% \begin{center}
%     \begin{eqnarray}
%         y^*   & = & \argmin_{y} \sum_{ i<j \in V } w_{ij} \cdot y_{ij} \\
%         s.t.: &  & y_{ij} + y_{jk} < y_{i,k} \quad \forall i, j, k   \nonumber
%     \end{eqnarray}
% \end{center}

% \paragraph{Node Coloring:}
% \begin{center}
%     \begin{eqnarray}
%         l^* &=& \argmin_{L} \sum_{ e_{ij} \in E } w_{ij} \cdot [l_{u} \neq l_{v}] \\
%         y_{ij}^* &=& [l_{u} \neq l_{v}]  
%     \end{eqnarray}
% \end{center}

@@

\section{Energy Based Hierarchical Clustering}\label{sec:ehc}
Agglomerative hierarchical clustering (HC) is widely used
in graph / image segmentation~\cite{arbelaez_2006}.
In each step, the edge with the highest weight $w$ is
contracted (red edges in Fig.~\ref{fig:hc_alg}).
Doing so, parallel edges edges can occur.
In agglomerative clustering, weights of
parallel edges are merged into single edges.
For image segmentation, the length
weighted mean is used to do this update~\cite{arbelaez_2006}.

While we, contrary to~\cite{arbelaez_2006}, directly work on energies, we are interesting in an 
agglomeration with respect to our energy function.
For this we propose to use energy based agglomeration
withe the following update rule:
Whenever there are multiple edges between 
a pair of nodes, these edges are merged into a single edge
by summing up their weights, since we 
minimize the sum of the cut edges.
We call HC with this update method Energy base Hierarchical Clustering (EHC).

We stop EHC if the highest edge weight is
smaller or equal to zero  (green edge in Fig.~\ref{fig:hc_alg}). Any further edge
contraction leads to worse energies.

% Hierarchical clustering is fast, and 
% due to the agglomeration of edge weights, 
% {\color{red} hierarchical clustering is robust against noise}.
% {\color{yellow} [ First most similar nodes are merged...]
% In agglomerative clustering, weights of
% parallel edges are merged into single edges.}
% For image segmentation, the length
% weighted mean is used to do this update [ref].

% [We contrary on those works work directly on energy functions and that why ...]
% We propose to use Energy Based Hierarchical Clustering (EHC) to find
% solutions for correlation clustering.

% To make EHC follow the energy, we propose 
% the following update rule:
% Whenever there are multiple edges between 
% a pair of nodes, these edges are merged into a single edge
% by summing up their weights, since we 
% minimize the sum of the cut edges.
% The algorithm stops if the highest edge weight is
% smaller or equal to zero  (green edge in fig. ~\ref{fig:hc_alg}). Any further edge
% contraction leads to worse energies.

Given the intrinsic greediness of hierarchical clustering, 
and the fact that EHC does only local decisions, it is 
sensitive against salt and paper noise 
and we do not expect very good solutions.

But EHC is very fast, and can be used to initialize
CGC~\cite{beier_2014_cvpr}. Excessive time in CGC
is spend in the \emph{cut phase} to solve the 
first 2 coloring on the complete graph.
As shown in Sec.~\ref{sec:exp},
allowing CGC to start from the EHC solution
can improve performance drastically.

% Common cluster distances
% \begin{itemize}
% \item Single-Link: Nearest Neighbor - their closest members.
% \item Complete-Link: Furthest Neighbor - their furthest members.
% \item Centroid: Distance between centroids
% \item Group Average: average of all cross-cluster pairs.
% \end{itemize}

%------------------------------------------------------------------------
\section{Correlation-Clustering-Fusion Moves}\label{sec:cc_fm}

Fusion moves as defined in \cite{Lempitsky-2010} work
in the node domain and do not work properly for 
objective functions as Eq.~\ref{eq:nodeproblem} since
the node coloring is ambiguous and has no semantic meaning (see Fig.~\ref{fig:fusion_abiguty}).
In the following, we propose a more suitable fusion move for correlation
clustering which works on the edge domain.
%
Given two proposal solutions $y'$ and $y''$,
$\mathcal{C}$ is the set of edges
which are uncut in $y'$ and $y''$.
%
\begin{align}
\breve{y}    & =  y' \cup y'' \\
\mathcal{C}  & =  \{ (i, j) \quad | \quad \breve{y}_{ij} = 0 \}
\end{align}
%
The fusion move for correlation clustering is solving Eq.~\ref{eq:edgeproblem}
with additional \emph{must-link constraints} for all edges in $\mathcal{C}$.
%
\begin{align}
  y^* &=& \argmin_{y \in P(G)} \sum_{ (i,j) \in E } w_{ij} \cdot y_{ij} \label{eq:fusion_move_a}.\\ 
  s.t & &  y_{ij} = 0 \quad \forall (i, j) \in \mathcal{C} \nonumber
\end{align}
%
By construction, solving Eq.~\ref{eq:fusion_move_a} cannot increase the energy:
\begin{align}
  \sum  w_{ij} \cdot y^*_{ij}  \leq \min \left( \sum  w_{ij} \cdot y'_{ij} ,  \sum   w_{ij} \cdot y''_{ij} \right)
\end{align}
%

Once we know how to fuse segmentations, the actual algorithm 
for correlation clustering 
is simple and similar to the fusion move algorithm of Lempitsky \etal~\cite{bagon_2011_arxiv}.
We iteratively improve the best solution by
fusing it with proposal solutions.
The inherent difference is how we define the fusion.
We fuse the segmentations according to Eq.~\ref{eq:fusion_move_a}.
%
How to optimize  Eq.~\ref{eq:fusion_move_a} in an more efficient way is
described in the following section.

\input{inputs/fig_fusion.tex}

%\textbf{old}
%
%Motivation Using Fusion moves -> problem ambiguity
%
%A consistent edge labeling $y$ defines a segmentation of $V$ by enumerating the connected components 
%in $G|_y=(V,\{e\in E|y_e=0\}$ by $l_v = \#(v)$, where $\#(v)$ denotes the number of the connected component of the %node $v\in V$.
%
%We define the contradiction of graph $G_y=(V_y,E_y)$
%with $V_y=\{\#(v)|v\in V\}$ and $E_y=\{\#(v),\#(u)|uv\in E, y_{uv}=1\})$.

%\input{inputs/fig_alg_graph.tex}

\subsection{Fast Optimization} \label{sec:fast_optimization}

Eq.~\ref{eq:fusion_move_a} can be solved efficiently, since 
we can work on a coarser graph, where all nodes which are connected
via must-link constraints are merged into single nodes.
This coarser graph is constructed by contracting all must-link edges
(see Fig.~\ref{fig:move_graph}).

TODO:
Here we need to proof that it is equivalent 
to optimize the contracted graph !

\input{inputs/fig_contraction.tex}
\input{inputs/fig_projection.tex}

\input{inputs/fig_move_graph.tex}

%-------------------------------------------------------------------------
%\subsection{Fusion Move Solver}

%-------------------------------------------------------------------------
\subsection{Proposal Generators}

>>>>

\section{Introduction}
Correlation clustering~\cite{Bansal-2002} also known as the multicut problem~\cite{chopra_1993_mp} 
is a basic primitive in computer vision~\cite{andres_2011_iccv,kroeger_2012_eccv,yarkony_2012_eccv,alush_2013_simbad} and data mining~\cite{Chierichetti-2014,Arasu-2009,Sadikov-2010,Chen-2012},
see Sec.~\ref{sec:problem_formulation} for its formal definition of clustering the nodes of a graph..

Its value is, firstly, that it accommodates both positive (attractive) \emph{and} negative (repulsive) edge weights,
which allows doing justice to evidence in the data that two nodes or pixels do \emph{not} wish  or do wish to end up in the same cluster or segment, respectively.
Secondly, it does not require a specification of the number of clusters beforehand.

Despite the clear mathematical formulation and nice properties,
correlation clustering is known to be NP-hard. 
%
Consequently, partition problems on large scale data, \eg
huge volume images in computational neuroscience~\cite{kroeger_2012_eccv}
or social networks~\cite{Leskovec-2010}, 
are not tractable, because reasonable solutions cannot be computed in acceptable time.

% Importantly, this allows doing justice to evidence in the data that two nodes or pixels do \emph{not} wish to end up in the same cluster or segment. This is in contrast to the submodular potentials so popularized by the graph cut algorithm, which only allow two nodes to be attracted to each other, or at most be agnostic about membership in the same cluster. Secondly, the algorithm does not require a specification of the number of clusters. This is in contrast to methods such as normalized cut \cite{} that can only accommodate attractive interactions and hence need the number of clusters to be specified.

% \paragraph{Contribution.} The evident usefulness of correlation clustering and its clean and compact formulation in terms of an optimization problem, together with its unfortunate NP-hardness, are an invitation to develop fast approximate solvers. The basic idea of the move making algorithm proposed here is to maintain, at all times, a best current partitioning; and to iteratively improve it (as we show, monotonously) by considering diverse and cheaply generated proposal partitionings. Any two nodes that are in one cluster in both the current and the proposal partitioning are contracted. Edges to contracted nodes are equally combined and reweighted, and the correlation clustering problem is then solved on this reduced graph. We offer a polyhedral characterization of this strategy, and evaluate it in conjunction with two versatile proposal generators.   We conduct experiments on a broad range of data sets ranging from 2D and 3D segmentation problems to clustering in signed social networks. The results suggest that this simple move making algorithm is the fastest technique known today, reaching close to globally optimal solutions in a tenth or a hundredth of the time required by the most efficient exact solvers. 
\vspace{0.1cm}
\noindent \textbf{Contribution.}
In the present work we present some novel approaches that are designed for large scale correlation clustering problems.
First, we define a novel energy based agglomerative clustering algorithm that monotonically increase the energy.
With this at hand we show how to improve the anytime performance of Cut, Clue \& Cut.
Second, we introduce cluster-fusion moves, which extend the original fusion moves~\cite{Lempitsky-2010} 
used in supervised segmentation to the unsupervised case and give a polyhedral interpretation of this algorithm.
We propose two versatile proposal generators, and evaluate the proposed methods on existing and new benchmark problems.
That experiments show that we can improve the computation time by one to two magnitudes without worsening the segmentation 
quality significantly.

\vspace{0.1cm}
\noindent \textbf{Related Work.}
A natural approach is to solve the integer linear programming problem directly. To this end, efficient separation procedures have been found \cite{} that allow to iteratively augment the set of constraints until a valid partitioning is found. 
Alternatively, it is possible to relax the integrality constraints of the integer linear programming formulation. Such an outer relaxation can be iteratively tightened. However, intermediate solutions are fractional and so rounding is required to obtain a valid partitioning. The latter approach works best on planar \cite{yarkony_2012_eccv} or almost planar \cite{yarkony-andres-2013} graphs.

@@

\vspace{0.1cm}
\noindent \textbf{Outline:} {\bf If we are short of space, the following can be ommited:}
In sec.~\ref{sec:problem_formulation} we will give a 
detailed problem definition where we introduce 
the correlation clustering objective.
In sec.~\ref{sec:related_work} we will 
discuss existing methods for correlation 
clustering and briefly explain the concept of fusion moves.
In sec: ~\ref{sec:cc_fm} we describe our proposed
method and show the properties of the algorithm.
In sec.~\ref{sec:exp} we show an evaluation
of the method on existing parameter  and discuss the effects of parameters.
Future work will be discussed in sec. \ref{sec:future} and
we will conclude in ~\ref{sec:conclusion}.
%-------------------------------------------------------------------------

\section{Notation and Problem Formulation}\label{sec:problem_formulation}
Let $G=(V,E, w)$ be a weighted graph of nodes $V$ and edges $E$.
%
The function $w : E \rightarrow \mathbb{R}$ assigns a weight to each edge.
A positive weight expresses the desire that two adjacent nodes should
be merged, whereas a negative weight indicates
that these nodes should be separated into two different regions.
%
%A \emph{subgraph} $G_A = \{A, E_A, w\}$ consists
%of nodes $A \subseteq V$ and edges $E_A := E\cap (A\times A)$.
%
A segmentation of the graph $G$ can by either given by an 
node labeling $l \in \mathbb{N}^{|V|}$
or an edge labeling $y \in\{0,1\}^{|E|}$.  
An edge labeling is only consistent if it contains no dangling edges~\cite{kappes_2013_arxiv}.
We denote the set of all consistent edge labelings by $P(G)\subset\{0,1\}^{|E|}$.
The convex hull of this set is known as the \emph{multicut polytope} $MC(G) = \textrm{conv}(P(G))$.

Given a weighted graph $G=(V,E,w)$ we consider the problem of segmenting $G$ such that the costs
of the edges between distinct segments is minimized. This can be formulated in the node domain
by assigning each node $v$ a label $l_v \in \mathbb{N}$
\begin{align}
  l^* &= \argmin_{L \in \mathbb{N}^{|V|}} \sum_{ (i,j) \in E } w_{ij} \cdot [l_{i} \neq l_{j}], \label{eq:nodeproblem}
\end{align} 
or in the edge domain, by label each edge $e$ as cut $y_e=1$ or uncut $y_e=0$ 
\begin{align}
  y^* &= \argmin_{y \in P(G)} \sum_{ (i,j) \in E } w_{ij} \cdot y_{ij} \label{eq:edgeproblem}.%\\ 
\end{align}
As shown in~\cite{kappes_2013_arxiv} both problems are equivalent, but formulation \ref{eq:nodeproblem}
suffers from ambiguities in the formulation~\cite{kappes_2011_emmcvpr}.

@@
